---
title: "Uniqueness of columns in tbl_df using dplyr"
output:
  html_document:
    df_print: paged
---

This document shows certain experiments with finding unique columns in a tbl_df using dplyr. The material is copied from https://stackoverflow.com/questions/22959635/remove-duplicated-rows-using-dplyr. 

We start by generating a random dataframe samples from a given distribution.

```{r}
suppressPackageStartupMessages(library(dplyr))
set.seed(123)
df <- data.frame(
  x = sample(0:1, 10, replace = T),
  y = sample(0:1, 10, replace = T),
  z = 1:10
)
df
```

Then we just want to keep the rows with unique entries over the columns `x` and `y`.

```{r}
(df %>% distinct(x, y))
```

As shown in the result, just columns `x` and `y` are kept. With the additional option `.keep_all = TRUE`, we get a different behavior.

```{r}
(df %>% distinct(x, y, .keep_all = TRUE))
```

We try to find which combinations of the grouping variables `x` and `y` occur how many times. 

```{r}
(df %>% group_by(x, y) %>% summarise(n = n()))
```

In case, we just want to find those which occur more than once, i.e., which are not unique, we filter on $n$. 

```{r}
(df %>% group_by(x, y) %>% summarise(n = n()) %>% filter(n > 1))
```

Another question is whether the whole thing also works with indices instead of names.

```{r}
(df %>% group_by(.[[1]], .[[2]]) %>% summarise(n = n()) %>% filter(n > 1))
```
