---
title: "dplyr - Testing New Verbs"
output: html_notebook
---

This notebook shows some scratch ideas and experiments with dplyr and the verbs to change certain values within a variable

## A simple example using case_when
We get started with a few examples from `?dplyr::case_when` 

```{r}
require(dplyr)
x <- 1:50
case_when(
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz",
  TRUE ~ as.character(x)
)
```

Doing the same thing with a replacement closer to our needs

```{r}
x <- 1:50
case_when(
  x %% 9  == 0 ~ "NA",
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz",
  TRUE ~ as.character(x)
)

```

Apparantly, we cannot use the "NA" without quotes. Trying with the typed version of NA

```{r}
x <- 1:50
case_when(
  x %% 9  == 0 ~ NA_character_,
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz" ,
  TRUE ~ as.character(x)
)

```



## Usage and Arguments
```{r, include=FALSE}
s_dpcw <- "`dplyr::case_when`"
s_dpcw_hf <- paste0("?", s_dpcw) 
```

From the sections __Usage__ and __Arguments__ in the help file of `r s_dpcw_hf`, we can see that `r s_dpcw` can take any number of arguments which is denoted by the three dots. The arguments must be a sequence of two sided formulas. A two-sided formula consists of a LHS and a RHS which are separated by a tilde (`~`). For the purpose of the arguments in `r s_dpcw` 


## Vector function
We are looking at the section __Vector Functions__ in a [RStudio blog post](https://blog.rstudio.com/2016/06/27/dplyr-0-5-0/). The second function in the section on __Vector Functions__ is called `na_if()`. The example used in the blog post is as follows 


```{r}
x <- c(1, 5, 2, -99, -99, 10)
na_if(x, -99)
```

The function `na_if()` is certainly useful when we want to replace a single value by NA. But we want to use it for a vector of values leading to

```{r}
na_if(x, c(1,2))
```

But the above produces an error. Hence we are searching for a different function. One possible candidate is `if_else()`. The example given in the blog post is

```{r}
x <- factor(sample(letters[1:5], 10, replace = TRUE))
if_else(x %in% c("a", "b", "c"), x, factor(NA))
```

Here an intersting side note is the difference to `ifelse()` in R-base

```{r}
ifelse(x %in% c("a", "b", "c"), x, factor(NA))
```

In the version of R-base, the components of `x` are replaced by the levels and are  no longer of type factor.












