---
title: "How to use dplyr Joins"
output:
  html_document:
    df_print: paged
---

This document examines how to join ages of parents to a record given for a certain animal.

## dplyr
We use the `dplyr`-vignette to learn how the different joins work.

### nycflights13 data
```{r}
### # check whether packages are installed
vec_pkgs <- c("magrittr", "nycflights13")
for (p in vec_pkgs){
  if (!p %in% installed.packages())
    install.packages(pkgs = p)
}

```

```{r}
library("nycflights13")
library(magrittr)
library(dplyr)

### # head of airlines dataset
head(airlines)
# Drop unimportant variables so it's easier to understand the join results.
flights2 <- flights %>% select(year:day, hour, origin, dest, tailnum, carrier)

flights2 %>% 
  left_join(airlines)
```


## Different Joins
Four different joins can be distinguished

```{r}
(df1 <- data_frame(x = c(1:3), y = 3:1))
(df2 <- data_frame(x = c(1, 3, 5), a = c(10,20,30), b = c("a","x","y")))
```

* `inner_join(x,y)`
```{r}
df1 %>% inner_join(df2) %>% knitr::kable()
```

* `left_join(x,y)`
```{r}
df1 %>% left_join(df2)
```

* `right_join(x,y)`
```{r}
df1 %>% right_join(df2)
```

* `full_join(x,y)`
```{r}
df1 %>% full_join(df2)
```


## Use Joins for pedigree checks
When, we want to check that mothers are older than their offspring, we join the birthdate of the mothers to the offspring records and check whether the consistence requirement is fullfilled.

First, we read the pedigree

```{r}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_final <- PedigreeFromTvdData::build_check_pedigree_from_tvd(ps_tvd_file = sDataFileName)
```

Get the column indices of the IDs as a list and column index of birthdates

```{r}
l_id_col_dsch_idx <- PedigreeFromTvdData::getTvdIdColsDsch()
n_bd_col_idx <- PedigreeFromTvdData::getBirthdateColIdxDsch()
```

Animal-Ids, Mother-Ids and birthdates are extracted from `tbl_ped_final`.

```{r}
tbl_animal <- tbl_ped_final %>% select(l_id_col_dsch_idx$TierIdCol, n_bd_col_idx, l_id_col_dsch_idx$MutterIdCol, l_id_col_dsch_idx$VaterIdCol)
names(tbl_animal) <- c("Tier", "Kalbedatum", "Mutter", "Vater")
head(tbl_animal)
```

Find animals that are also mothers

```{r}
tbl_mother_ids <- tbl_animal %>% filter(Mutter != "") %>% select(Mutter)
head(tbl_mother_ids)
nrow(tbl_mother_ids)
```

Join ids of mother to the animal records to obtain the birthdate of the mothers

```{r}
tbl_mother_ids %>% inner_join(tbl_animal, by = c("Mutter" = "Tier"))
(tbl_mother_bd <- tbl_mother_ids %>% inner_join(tbl_animal, by = c("Mutter" = "Tier")) %>% select(Mutter, Kalbedatum))
```

Join the mothers and their birthdates back to the animal records

```{r}
(tbl_ani_mother_bd <- tbl_mother_bd %>% inner_join(tbl_animal, by = "Mutter"))

```

Filter according to the consistency requirements

```{r}
tbl_ani_mother_bd %>% filter(Kalbedatum.x > Kalbedatum.y)
tbl_ani_mother_bd %>% filter((Kalbedatum.y - Kalbedatum.x) < 1000)

```

