---
title: "Pedigree Checks - Implementations"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pedigree Checks - Implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)

```

## Disclaimer
This vignette describes implementation of checks that are described in the companion vignette on strategies entitled [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). 

## Checks
The vignette on the strategy defined checks to be routines that search for pedigree records that do not fullfill certain requirement conditions. The different type of checks have different requirements in computing resources. The check whether the pedigree contains any cycles is quite expensive to compute. One possible way to implement this check is to use an algorithm that computes the toplogical sort of the pedigree nodes. This construction does not give a result, whenever the pedigree contains any loops. For that reason this check is currently not implemented. 

### Checking for unique IDs
As the result returned from checks are lists of primary keys of pedigree records, we must ensure that these primary keys are available. This must be done right after reading the pedigree from the input file.

The pedigree is read using the following statement

```{r}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_uni_id <- laf_open_fwf_tvd_input(ps_input_file = sDataFileName, pb_out = TRUE)
```

As mentioned above, we first check for uniqueness of the animal-IDs. This is done with the following function call

```{r}
tbl_non_uni <- check_unique_animal_id(tbl_ped_uni_id, pb_out = TRUE)
# check whether there are any non-unique ids
if (nrow(tbl_non_uni) > 0){
  cat(" *** Nr of duplicate IDs found: ", nrow(tbl_non_uni), "\n")
} else {
  cat(" *** No duplicate IDs found: ", nrow(tbl_non_uni), "\n")
}
```

For a more detailed check of the function, we have created a data set that contains duplicate IDs. This constructed dataset is now used for a further test.

```{r}
s_data_file_dup_id <- system.file(file.path("extdata","KLDAT_20171214_dup_id_100.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_dup_id <- laf_open_fwf_tvd_input(ps_input_file = s_data_file_dup_id, pb_out = TRUE)
(tbl_non_uni_dup_id <- check_unique_animal_id(tbl_ped_dup_id, pb_out = TRUE))
```

If you want to know which records are duplicate then
```{r}
(tbl_ped_dup_id %>% filter(V12==tbl_non_uni_dup_id$Animal[1]))
```


Whenever, the check for uniqueness of the IDs is successful the property of the maximum in-degree of every node is also fullfilled. Hence we do not need to write a separate function for a check on the maximum in-degree.

### Parents older than offspring
The check whether parents are older than offspring is done with the function `check_parent_older_offspring()` which has to be called for dams and sires separately. Per default the difference of age between parents and offspring have to be bigger than 1 year `pn_date_diff_tol = 10^4`.

For the morthers

```{r}
(l_tvd_id_col_dsch <- getTvdIdColsDsch())
(n_bd_col_idx <- getBirthdateColIdxDsch())
(tbl_fail_mother_age <- check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id, 
                                         pn_offspring_col = l_tvd_id_col_dsch$TierIdCol, 
                                         pn_birthday_col = n_bd_col_idx, 
                                         pn_parent_col = l_tvd_id_col_dsch$MutterIdCol))
```

The same for the fathers

```{r}
(tbl_fail_father_age <- check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id, 
                                         pn_offspring_col = l_tvd_id_col_dsch$TierIdCol, 
                                         pn_birthday_col = n_bd_col_idx, 
                                         pn_parent_col = l_tvd_id_col_dsch$VaterIdCol))
```

### Parents sex consistency
The check if the sex of the parents are consistent is done with the function `check_sex_tbl()` which has to be called for dams and sires separately. 

For mothers and fathers:
```{r}
(tbl_result_consistencysex <- check_sex_tbl(ptblPedigree = tbl_ped_uni_id))
```

We have created a data set that contains mother with inconsistent sex. This constructed dataset is now used for a further test
```{r}
sDataFileSex <- system.file(file.path("extdata","KLDAT_20170524_100_mother_sex.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_sex <- laf_open_fwf_tvd_input(ps_input_file = sDataFileSex, pb_out = TRUE)
(tbl_result_sex <- check_sex_tbl(ptblPedigree = tbl_ped_sex))
```

###Correct format of IDs for individuals and parents
The format of TVD-IDs will be checked: position 1 to 2 have to be letters and the rest numbers.
This function hast to be call each time to check tvdid of individual, mother and father.

Here is an exemple with the check for individual:

```{r}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"),package = "PedigreeFromTvdData")
tbl_ped_uni_id <- laf_open_fwf_tvd_input(ps_input_file = sDataFileName, pb_out = TRUE)
tbl_ped_uni_id <- tbl_ped_uni_id[c(1:100),]

lIdCols <- getTvdIdColsDsch()
(tbl_result_tvdformat <- correct_tvd_format_tbl(p_tbl_ped = tbl_ped_uni_id,
                                   plFormatBorder = getTVDIdBorder(),
                                   plIdCols = lIdCols,
                                   pnIdCol = lIdCols$TierIdCol))
```

We have created a data set that contains inconsistent IDs format. This constructed dataset is now used for a further test
```{r}
sData_NotCorrectId <- system.file(file.path("extdata","KLDAT_20170524_20_notcorrectId.txt"),
                                  package ="PedigreeFromTvdData")
tbl_ped_correctFormat <- laf_open_fwf_tvd_input(ps_input_file = sData_NotCorrectId, pb_out = TRUE)
lIdCols <- getTvdIdColsDsch()
(tbl_result_tvdformat <- correct_tvd_format_tbl(p_tbl_ped = tbl_ped_correctFormat,
                                   plFormatBorder = getTVDIdBorder(),
                                   plIdCols = lIdCols,
                                   pnIdCol = lIdCols$TierIdCol))

```

TO DO: Check influence of too short Ids or records!

###Correct format of bithdates
The format of birthdates will be checked: days (1-31), months (1-12), years (>1950)

```{r}
(tbl_result_checkbirthdate <- check_birthdate_tbl(ptblPedigree = tbl_ped_uni_id))
```

We have created a data set that contains inconsistent Birthdates. This constructed dataset is now used for a further test

```{r}
sData_NotCorrectBirthdate <- system.file(file.path("extdata","KLDAT_20170524_20_notcorrectBirthdate.txt"),
                                  package ="PedigreeFromTvdData")
tbl_ped_NcorrectBirthdate <- laf_open_fwf_tvd_input(ps_input_file = sData_NotCorrectBirthdate, pb_out = TRUE)

(tbl_result_checkbirthdate_nC <- check_birthdate_tbl(ptblPedigree = tbl_ped_NcorrectBirthdate))

```


##Test with a big data set
Checking for unique IDs
```{r}
(mb.uniqId <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::check_unique_animal_id(ptbl_pedigree = tbl_ped_uni_id),
  times = 5, unit = "s"))
```

Parents older than offspring
```{r}
l_tvd_id_col_dsch <- getTvdIdColsDsch()
n_bd_col_idx <- getBirthdateColIdxDsch()
(mb.parentOld <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id,
                                                                pn_offspring_col = l_tvd_id_col_dsch$TierIdCol, 
                                                                pn_birthday_col = n_bd_col_idx, 
                                                                pn_parent_col = l_tvd_id_col_dsch$MutterIdCol),
  times = 5, unit = "s"))
```

Parents sex consistency
```{r}
(mb.sexConsistency <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::check_sex_tbl(ptblPedigree = tbl_ped_uni_id),
  times = 5, unit = "s"))
```

Correct format of IDs for individuals
```{r}
lIdCols <- getTvdIdColsDsch()
(mb.format <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::correct_tvd_format_tbl(p_tbl_ped = tbl_ped_uni_id,
                                                          pnIdCol = lIdCols$TierIdCol),
  times = 5, unit = "s"))
```

Correct format of bithdates
```{r}
(mb.birthdate <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::check_birthdate_tbl(ptblPedigree = tbl_ped_uni_id),
  times = 5, unit = "s"))
```

Overview of results
The following table shows the median times (column MedianTime) in seconds for all function that were used.
```{r MedTimeRead, echo=FALSE}
vecMethods <- c("PedigreeFromTvdData::check_unique_animal_id",
                "PedigreeFromTvdData::check_parent_older_offspring",
                "PedigreeFromTvdData::check_sex_tbl",
                "PedigreeFromTvdData::correct_tvd_format_tbl",
                "PedigreeFromTvdData::check_birthdate_tbl")
vecMedTimes <- c(mb.uniqId$time[3],
                 mb.parentOld$time[3],
                 mb.sexConsistency$time[3],
                 mb.format$time[3],
                 mb.birthdate$time[3])
nMinTime <- vecMedTimes[order(vecMedTimes)][1]
dfMedTime <- data.frame(Methode = vecMethods,
                        MedianTime = round(vecMedTimes*10^(-9), digits = 4))
knitr::kable(dfMedTime)
```


## Session Info
```{r}
sessionInfo()
```


## Latest Update
`r paste(Sys.time(),paste0("(", Sys.info()[["user"]],")" ))`

