---
title: "Main Pedigree Build"
author: "Peter von Rohr and Sophie Kunz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main Pedigree Build}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)
```


## Disclaimer
This document describes a function implementing different functions to check and transform a pedigree. A general description of the strategies is available in a separate vignette on [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). Further description about the consistency checks implementation is available on [Pedigree Checks - Implementations](ConsistencyCheckImplementation.html) and a description about the transformation implementation is available on [Pedigree Transformation - Implementations](TransformationImplementation.html).

```{r}
sDataFileErrors <- system.file(file.path("extdata","KLDAT_20170524_AllFehler.txt"), 
                             package = "PedigreeFromTvdData")

(tbl_transform_ped <- build_check_pedigree_from_tvd(ps_tvd_file = sDataFileErrors,
                              pvec_format = getK11ColPositionVecFromDsch()))
```

Remaining:
All dupicate are deleted;
Parent and offspring birthdate are set to NA, if a parent is not older as offspring;
Animal sex is set to NA, if the sex is incorrect;
Birthdate is set to NA, if the birthdate format is incorrect.

Time how long is this function running.
```{r}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")

(mb.main <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::build_check_pedigree_from_tvd(ps_tvd_file = sDataFileName,
                                                                 pvec_format = getK11ColPositionVecFromDsch()),
  times = 5, unit = "s"))
```

Overview of results
The following table shows the median times (column MedianTime) in seconds for the function that was used.
```{r MedTimeRead, echo=FALSE}
vecMethods <- c("PedigreeFromTvdData::build_check_pedigree_from_tvd")
vecMedTimes <- c(mb.main$time[3])
nMinTime <- vecMedTimes[order(vecMedTimes)][1]
dfMedTime <- data.frame(Methode = vecMethods,
                        MedianTime = round(vecMedTimes*10^(-9), digits = 4))
knitr::kable(dfMedTime)
```

## Session Info
```{r}
sessionInfo()
```


## Latest Update
`r paste(Sys.time(),paste0("(", Sys.info()[["user"]],")" ))`
