---
title: "CheckList for TODOs"
author: "Sophie Kunz und Peter von Rohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CheckList for TODOs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Disclaimer
This document contains issues and problems that are to be solved at a later point in time. Whenever a solution is proposed, the solution can be included into this document.


## Data Format of Birthdates
Currently, dates are stored as integer values. This makes the comparison between dates such as birthdates of parents and offspring rather complicated an inflexible.

A proposed solution is to store dates as date values. See the chapter on [Dates and Times](http://r4ds.had.co.nz/dates-and-times.html) in R for Data Science.

## Breed
Breed in TierRassecode has empty space at the end. Some Rassecode have 2 letters and other 3.
Further by merging, a function has to be implement in order that TVD-Rassecode for a breed equal to the one of Qualitas. Christian Stricker has in the routine evaluation implemented under 'verifyBreed'.


## Session Info
```{r}
sessionInfo()
```


## Latest Update
`r paste(Sys.time(),paste0("(", Sys.info()[["user"]],")" ))`
