---
title: "Reading Fixed Width Formatted Data"
author: "Sophie Kunz and Peter von Rohr"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Reading Fixed Width Formatted Data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r PackageSetup, eval=FALSE, include=FALSE}
for (pgk in c("readr", "iotools", "LaF", "microbenchmark")){
  if (require(pgk, character.only = TRUE))
    install.packages(pkgs = pgk)
}
```

# Disclaimer
This vignette compares different methods of reading large amounts of fixed width formatted (fwf) data from a file. 


# Introduction
We are given a large file (about 13 million lines and 1.3 Gb of data) with fixed formatted data. This data needs to be read into R to do some consistency checks on the data and to write the data to a file in a different format. In a first implementation, we used the function `readLines()` to read each line, parse the content and store the data in a nested list. 

## Scaling
This approach worked well for small datasets, but it does not scale to bigger sizes of the input data. Even after several days of wall clock time, the data is still no completely processed. Hence we need a faster method to read the large input file and to process the data.

## Remainder
In this document, we are comparing different methods for reading large amounts of data. The focus will be on R-packages `iotools` and `LaF`. Because the methods of how the data will be processed is not completely independent of how the data is read from the file, we are going to discuss different approaches on how the data can be processed towards the end of this document.


# Methods
We are starting by comparing different methods of how we can read large amounts of data into R. 

## readlines
The original idea that has shown to be very slow was to read the input line-by-line using `readline()` and building the required data-structure for consistency checking as nested list. 

## read.fwf
The standard way of reading fwf-data into an R dataframe is to use the function read.fwf from the base package.

## read_fwf
The package `readr` is part of the `tidyverse` and it offers the function `read_fwf` to read fwf-data. According to https://github.com/tidyverse/readr the function `data.table::fread()` is even faster than `readr::read.fwf()`

## iotools
The package `iotools` is available from http://www.rforge.net/iotools/ and https://cran.r-project.org/web/packages/iotools/index.html and described in https://arxiv.org/abs/1510.00041. 

## LaF
The package `LaF` is available from https://cran.r-project.org/web/packages/LaF/index.html and https://github.com/djvanderlaan/LaF. This package allows to read data from files that are too big to fit into memory which is a special property that is not available in other tools.



# Results
We are testing the described tools using a subset of the complete input data. This should make the tests to be runnable in an interactive mode.

```{r}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")
vecFormat <- c(22,14,3,31,8,14,3,1,14,3)
```

The size of this file in MB is 

```{r}
round(file.info(sDataFileName)$size / 1024 / 1024, digits = 2)
```

This file is read using each of the above mentioned methods and the time it takes to read them is compared. 

```{r}
mb <- microbenchmark::microbenchmark(
  data.fwf = read.fwf(file = sDataFileName, widths = vecFormat),
  data_fwf = suppressMessages(readr::read_fwf(file = sDataFileName, 
                              col_positions = readr::fwf_widths(widths = vecFormat))),
  iotools  = iotools::dstrfw(iotools::readAsRaw(sDataFileName), 
                            col_types = rep("character", length(vecFormat)),
                            widths = vecFormat),
  laf = LaF::laf_open_fwf(filename = sDataFileName, 
                          column_types = rep("character", length(vecFormat)),
                          column_widths = vecFormat),
  times = 5
)
mb
```


# Resources
- Comparison of different methods for reading fwf data: 
[Faster way to read fixed-width files in R](https://stackoverflow.com/questions/24715894/faster-way-to-read-fixed-width-files-in-r)
- Alternative SO-post: [Reading big data with fixed width](https://stackoverflow.com/questions/18720036/reading-big-data-with-fixed-width)
- paper on R-package iotools: https://arxiv.org/pdf/1510.00041.pdf
- R-package LaF: https://cran.r-project.org/web/packages/LaF/vignettes/LaF-manual.pdf
