---
title: "Pedigree Transformation - Implementations"
author: "Sophie Kunz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pedigree Transformation - Implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)
```

## Disclaimer
This vignette describes implementation of transformation that are described in the companion vignette on strategies entitled [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). 

## Transformations
The vignette on the strategy defined transformations to be routines that search for pedigree records that do not fullfill certain requirement conditions. Transformations are either delete a complete pedigree record or invalidate with NA a field of the pedigree record.

### Tranformation after cheks of uniqueness of animal-IDs
Animal-IDs is the primary key. If duplicate for animal-ID have been found. The complete record has to be deleted.

```{r DuplicateId, eval=TRUE, include=FALSE, results='hide'}
s_data_file_dup_id <- system.file(file.path("extdata","KLDAT_20171214_dup_id_100.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_dup_id <- laf_open_fwf_tvd_input(ps_input_file = s_data_file_dup_id, pb_out = TRUE)
# (tbl_non_uni_dup_id <- check_unique_animal_id(tbl_ped_dup_id, pb_out = TRUE))
```


```{r}
(tbl_transform_ped <- transform_unique_animal_id(ptbl_pedigree = tbl_ped_dup_id,
                                                 pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol,
                                                 pb_out = TRUE))
```




```{r}
transform_unique_animal_id <- function(ptbl_pedigree,
                                       pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol,
                                       pb_out = FALSE){

  ### # assign result that will be returned
  tbl_transform_ped <- ptbl_pedigree

  ### # Output number of rows in original pedigree for debuggin
   if (pb_out) {
     cat(" *** Unique ID pedigree transformation on original pedigree with nr records: ",
         nrow(tbl_transform_ped), "\n")
     # cat(" *** Non-unique Ids found: ", ptbl_pedigree %>% filter(V12==output_check$Animal))
   }

  ### # check whether we have to do a transformation using the check function
  tbl_non_uni_dup_id <- check_unique_animal_id(ptbl_pedigree = tbl_transform_ped)
  ### # debugging output with number of duplicate records
  if (pb_out)
    cat(" *** Number of records with non unique ids found: ", nrow(tbl_non_uni_dup_id), "\n")
  ### # if duplicate records are found, do the transformation by retaining only
  ### #  the unique records.
  if (nrow(tbl_non_uni_dup_id) > 0) {
    tbl_transform_ped <- tbl_transform_ped %>%
      group_by(.[[pn_ani_id_col_idx]]) %>%
      summarise(n = n()) %>%
      filter((n == 1))
      #filter(!(n > 1)) #Sophie -> Alle Non-Unique Ids löschen, keine behalten
  }

  ### # debugging output with number of unique pedigree records after transformation
  if (pb_out){
    cat(" *** Number of records after transformation: ", nrow(tbl_transform_ped))
    # cat(" *** Transformation duplicate Ids, none should be found : ", tbl_transform_ped %>% filter(V12==output_check$Animal))
  }

  ### # return result
  return(tbl_transform_ped)
}

```



