---
title: "Pedigree Transformation - Implementations"
author: "Sophie Kunz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pedigree Transformation - Implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)
```

## Disclaimer
This vignette describes implementation of transformation that are described in the companion vignette on strategies entitled [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). 

## Transformations
The vignette on the strategy defined transformations to be routines that search for pedigree records that do not fullfill certain requirement conditions. Transformations are either delete a complete pedigree record or invalidate with NA a field of the pedigree record.

### Tranformation after cheks of uniqueness of animal-IDs
Animal-IDs is the primary key. If duplicate for animal-ID have been found. The complete record has to be deleted.

```{r DuplicateId, eval=TRUE, include=FALSE, results='hide'}
s_data_file_dup_id <- system.file(file.path("extdata","KLDAT_20171214_dup_id_100.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_dup_id <- laf_open_fwf_tvd_input(ps_input_file = s_data_file_dup_id, pb_out = TRUE)
# (tbl_non_uni_dup_id <- check_unique_animal_id(tbl_ped_dup_id, pb_out = TRUE))
```


```{r}
(tbl_transform_ped <- transform_unique_animal_id(ptbl_pedigree = tbl_ped_dup_id,
                                                 pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol,
                                                 pb_out = TRUE))
```

