---
title: "Pedigree Transformation - Implementations"
author: "Sophie Kunz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pedigree Transformation - Implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)
```

## Disclaimer
This vignette describes implementation of transformation that are described in the companion vignette on strategies entitled [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). 

## Transformations
The vignette on the strategy defined transformations to be routines that search for pedigree records that do not fullfill certain requirement conditions. Transformations are either delete a complete pedigree record or invalidate with NA a field of the pedigree record.

### Tranformation after checks of uniqueness of animal-IDs
Animal-IDs are the primary key. If duplicates of animal-IDs have been found, the complete record has to be deleted. Those records with duplicate animal-IDs which have more information in the other fields of the pedigree record are kept. If the pedigree records with the duplicate animal-IDs are identical the record that is kept is selected randomly.

```{r DuplicateId, eval=TRUE, include=FALSE, results='hide'}
s_data_file_dup_id <- system.file(file.path("extdata","KLDAT_20171214_dup_id_100.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_dup_id <- laf_open_fwf_tvd_input(ps_input_file = s_data_file_dup_id, pb_out = TRUE)
(tbl_non_uni_dup_id <- check_unique_animal_id(tbl_ped_dup_id, pb_out = TRUE))
```


```{r}
(tbl_transform_ped <- transform_unique_animal_id(ptbl_pedigree = tbl_ped_dup_id,
                                                 pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol,
                                                 pb_out = TRUE))
```

The above version of the transformation function `transform_unique_animal_id()` keeps only the records which have a unique ID. The records with duplicate IDs are discarded. It could well be possible that we want to keep one record of those with duplicate records. In that case, we certainly want to keep the one with the most information available.

To implement this approach of keeping one record of those with duplicate IDs, we have to inspect those records with duplicate IDs a bit closer. These records are found as a result of the check-function `check_unique_animal_id()`. We run a join to see what information is available from these records.

```{r}
tbl_ped_dup_id %>% inner_join(tbl_non_uni_dup_id, by = c("V12" = "Animal"))  %>% select(V5,V11,V12,V16)
```

The above shown output points to a problem that the information available from the records with duplicate IDs can be ambiguous. For example, birthdates can be different. Without information about the parents, we cannot say which birthdate is correct. The required information about the parents can only become available when we try to merge with the pedigree in ARGUS.


### Transformation after checks, if parents are older than offspring
If the parent are older than offspring, the parent's birthdate field will be invalidate with NA.

```{r ParentYounger, eval=TRUE, include=FALSE, results='hide'}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_uni_id <- laf_open_fwf_tvd_input(ps_input_file = sDataFileName, pb_out = TRUE)
(l_tvd_id_col_dsch <- getTvdIdColsDsch())
(n_bd_col_idx <- getBirthdateColIdxDsch())
(tbl_fail_mother_age <- check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id, 
                                         pn_offspring_col = l_tvd_id_col_dsch$TierIdCol, 
                                         pn_birthday_col = n_bd_col_idx, 
                                         pn_parent_col = l_tvd_id_col_dsch$MutterIdCol))
```

#### Single animal-IDs
The result of the above check give us a dataframe containing animals and parents which have wrong birthdates. According to https://stackoverflow.com/questions/27909000/set-certain-values-to-na-with-dplyr, we can use dplyr::mutate and replace. But this shows only how to do it with one ID and not how this could work with a vector of IDs. 

We start with a single ID

```{r}
### # select first animal-ID
(s_single_id <- tbl_fail_mother_age$Animal[1])
### # assign result dataframe and do the mutation
tbl_transform_ped <- tbl_ped_uni_id %>% mutate(V11 = replace(V11, which(V12 == s_single_id), NA))
```

Checking whether the birthdate of the pedigree record for animal with the ID `r s_single_id` was changed can be done by

```{r}
tbl_transform_ped %>% filter(V12 == s_single_id) %>% select(V5,V11,V12,V16)
```

#### Multiple animal-IDs
The above approach can be used, as soon as we can provide an index vector of all records of which the birthdate must be changed. 

```{r}
(vec_ani_ids <- c(tbl_fail_mother_age$Animal, tbl_fail_mother_age$Parent))
vec_ani_idx <- sapply(vec_ani_ids, function(x) which(tbl_ped_uni_id$V12 == x), USE.NAMES = FALSE)
tbl_transform_ped <- tbl_ped_uni_id %>% mutate(V11 = replace(V11, vec_ani_idx, NA))
```
Checking the records

```{r}
tbl_transform_ped %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Animal")) %>% select(V12,V11)
tbl_transform_ped %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Parent")) %>% select(V12,V11)
```

Comparing the transformed pedigree records to the original records shows that birthdates are invalidated.

```{r}
tbl_ped_uni_id %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Animal")) %>% select(V12,V11)
tbl_ped_uni_id %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Parent")) %>% select(V12,V11)
```

#### Function transform_check_parent_older_offspring

```{r}
(tbl_transform_ped <- transform_check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id,
                                                            output_check = tbl_fail_mother_age,
                                                            pb_out = TRUE))
```

### Tranformation by sex inconsistency
If the sex of the parent are inconsistent, the parent's sex field will be invalidate with NA.

```{r ParentSexInconsistency, eval=TRUE, include=FALSE, results='hide'}
sDataFileSex <- system.file(file.path("extdata","KLDAT_20170524_100_mother_sex.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_sex <- laf_open_fwf_tvd_input(ps_input_file = sDataFileSex, pb_out = TRUE)
tbl_result_sex <- check_sex_tbl(ptblPedigree = tbl_ped_sex)
```

```{r}
(tbl_transform_ped <- transform_check_sex_tbl(ptbl_pedigree = tbl_ped_sex,
                         output_check = tbl_result_sex,
                         pb_out = TRUE))
```

###Transformation by incorrect IDs format
If the format of TVD-IDs is not correct, the complete record of an animal has to be deleted. If the parent TVD-Ids is not correct, the field TVDids will be with NA invalidated.

```{r NotCorrectTVDFormat, eval=TRUE, include=FALSE, results='hide'}
sData_NotCorrectId <- system.file(file.path("extdata","KLDAT_20170524_20_notcorrectId.txt"),
                                  package ="PedigreeFromTvdData")
tbl_ped_correctFormat <- laf_open_fwf_tvd_input(ps_input_file = sData_NotCorrectId, pb_out = TRUE)
lIdCols <- getTvdIdColsDsch()
tbl_result_tvdformat <- correct_tvd_format_tbl(p_tbl_ped = tbl_ped_correctFormat,
                                   plFormatBorder = getTVDIdBorder(),
                                   plIdCols = lIdCols,
                                   pnIdCol = lIdCols$TierIdCol)
```

Vorsicht: Unterschied wenn das Format bei Tier oder Eltern nicht korrekt ist. Beim Tier das record löschen, bei den Eltern nur das Feld mit NA setzen. Jetzt wird nur das Feld TVD auf NA gesetzt.
```{r}
(tbl_transform_ped <- transform_correct_tvd_format_tbl(ptbl_pedigree = tbl_ped_correctFormat,
                                                       output_check = tbl_result_tvdformat,
                                                       pb_out = TRUE))
```

###Transformation by incorrect format of bithdates
If the format of birthdate is not correct, the birthdate field will be invalidate with NA.

```{r NotCorrectBirthdateFormat, eval=TRUE, include=FALSE, results='hide'}
sData_NotCorrectBirthdate <- system.file(file.path("extdata","KLDAT_20170524_20_notcorrectBirthdate.txt"),
                                  package ="PedigreeFromTvdData")
tbl_ped_NcorrectBirthdate <- laf_open_fwf_tvd_input(ps_input_file = sData_NotCorrectBirthdate, pb_out = TRUE)

tbl_result_checkbirthdate_nC <- check_birthdate_tbl(ptblPedigree = tbl_ped_NcorrectBirthdate)
```

```{r}
(tbl_transform_ped <- transform_check_birthdate_tbl(ptbl_pedigree = tbl_ped_NcorrectBirthdate,
                                                    output_check = tbl_result_checkbirthdate_nC,
                                                    pb_out = TRUE))
```

##Test with a big data set
```{r BigDat, eval=TRUE, include=FALSE, results='hide'}

```


Tranformation after checks of uniqueness of animal-IDs
```{r}
(mb.uniqId <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::transform_unique_animal_id(ptbl_pedigree = tbl_ped_uni_id,
                                                 pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol),
  times = 5, unit = "s"))
```

Transformation after checks, if parents are older than offspring
```{r}
(mb.parentOld <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::transform_check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id,
                                                                output_check = tbl_fail_mother_age),
  times = 5, unit = "s"))
```

Tranformation by sex inconsistency
```{r}
(mb.sexConsistency <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::transform_check_sex_tbl(ptbl_pedigree = tbl_ped_uni_id,
                                                           output_check = check_sex_tbl(ptblPedigree = tbl_ped_uni_id)),
  times = 5, unit = "s"))
```

Transformation by incorrect IDs format
```{r}
lIdCols <- getTvdIdColsDsch()
(mb.format <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::transform_correct_tvd_format_tbl(ptbl_pedigree = tbl_ped_uni_id,
                                                          output_check = correct_tvd_format_tbl(p_tbl_ped = tbl_ped_uni_id,
                                   plFormatBorder = getTVDIdBorder(),
                                   plIdCols = lIdCols,
                                   pnIdCol = lIdCols$TierIdCol)),
  times = 5, unit = "s"))
```

Transformation by incorrect format of bithdates
```{r}
(mb.birthdate <- microbenchmark::microbenchmark(
  data.base = PedigreeFromTvdData::transform_check_birthdate_tbl(ptbl_pedigree = tbl_ped_uni_id,
                                                                 output_check = check_birthdate_tbl(ptblPedigree = tbl_ped_uni_id)),
  times = 5, unit = "s"))
```

Overview of results
The following table shows the median times (column MedianTime) in seconds for all function that were used.
```{r MedTimeRead, echo=FALSE}
vecMethods <- c("PedigreeFromTvdData::transform_unique_animal_id",
                "PedigreeFromTvdData::transform_check_parent_older_offspring",
                "PedigreeFromTvdData::transform_check_sex_tbl",
                "PedigreeFromTvdData::transform_correct_tvd_format_tbl",
                "PedigreeFromTvdData::transform_check_birthdate_tbl")
vecMedTimes <- c(mb.uniqId$time[3],
                 mb.parentOld$time[3],
                 mb.sexConsistency$time[3],
                 mb.format$time[3],
                 mb.birthdate$time[3])
nMinTime <- vecMedTimes[order(vecMedTimes)][1]
dfMedTime <- data.frame(Methode = vecMethods,
                        MedianTime = round(vecMedTimes*10^(-9), digits = 4))
knitr::kable(dfMedTime)
```


## Session Info
```{r}
sessionInfo()
```


## Latest Update
`r paste(Sys.time(),paste0("(", Sys.info()[["user"]],")" ))`


