---
title: "Pedigree Transformation - Implementations"
author: "Sophie Kunz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pedigree Transformation - Implementations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r PackageSetup, eval=TRUE, include=FALSE, results='hide'}
require(PedigreeFromTvdData)
require(dplyr)
require(magrittr)
```

## Disclaimer
This vignette describes implementation of transformation that are described in the companion vignette on strategies entitled [Pedigree Checks - Concepts and Strategies](ConsistencyCheckStrategy.html). 

## Transformations
The vignette on the strategy defined transformations to be routines that search for pedigree records that do not fullfill certain requirement conditions. Transformations are either delete a complete pedigree record or invalidate with NA a field of the pedigree record.

### Tranformation after checks of uniqueness of animal-IDs
Animal-IDs are the primary key. If duplicates of animal-IDs have been found, the complete record has to be deleted. Those records with duplicate animal-IDs which have more information in the other fields of the pedigree record are kept. If the pedigree records with the duplicate animal-IDs are identical the record that is kept is selected randomly.

```{r DuplicateId, eval=TRUE, include=FALSE, results='hide'}
s_data_file_dup_id <- system.file(file.path("extdata","KLDAT_20171214_dup_id_100.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_dup_id <- laf_open_fwf_tvd_input(ps_input_file = s_data_file_dup_id, pb_out = TRUE)
# (tbl_non_uni_dup_id <- check_unique_animal_id(tbl_ped_dup_id, pb_out = TRUE))
```


```{r}
(tbl_transform_ped <- transform_unique_animal_id(ptbl_pedigree = tbl_ped_dup_id,
                                                 pn_ani_id_col_idx = getTvdIdColsDsch()$TierIdCol,
                                                 pb_out = TRUE))
```

### Transformation after checks, if parents are older than offspring
If the parent are older than offspring, the parent's birthdate field will be invalidate with NA.

```{r ParentYounger, eval=TRUE, include=FALSE, results='hide'}
sDataFileName <- system.file(file.path("extdata","KLDAT_20170524_10000.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_uni_id <- laf_open_fwf_tvd_input(ps_input_file = sDataFileName, pb_out = TRUE)
(l_tvd_id_col_dsch <- getTvdIdColsDsch())
(n_bd_col_idx <- getBirthdateColIdxDsch())
(tbl_fail_mother_age <- check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id, 
                                         pn_offspring_col = l_tvd_id_col_dsch$TierIdCol, 
                                         pn_birthday_col = n_bd_col_idx, 
                                         pn_parent_col = l_tvd_id_col_dsch$MutterIdCol))
```

#### Single animal-IDs
The result of the above check give us a dataframe containing animals and parents which have wrong birthdates. According to https://stackoverflow.com/questions/27909000/set-certain-values-to-na-with-dplyr, we can use dplyr::mutate and replace. But this shows only how to do it with one ID and not how this could work with a vector of IDs. 

We start with a single ID

```{r}
### # select first animal-ID
(s_single_id <- tbl_fail_mother_age$Animal[1])
### # assign result dataframe and do the mutation
tbl_transform_ped <- tbl_ped_uni_id %>% mutate(V11 = replace(V11, which(V12 == s_single_id), NA))
```

Checking whether the birthdate of the pedigree record for animal with the ID `r s_single_id` was changed can be done by

```{r}
tbl_transform_ped %>% filter(V12 == s_single_id) %>% select(V5,V11,V12,V16)
```

#### Multiple animal-IDs
The above approach can be used, as soon as we can provide an index vector of all records of which the birthdate must be changed. 

```{r}
(vec_ani_ids <- c(tbl_fail_mother_age$Animal, tbl_fail_mother_age$Parent))
vec_ani_idx <- sapply(vec_ani_ids, function(x) which(tbl_ped_uni_id$V12 == x), USE.NAMES = FALSE)
tbl_transform_ped <- tbl_ped_uni_id %>% mutate(V11 = replace(V11, vec_ani_idx, NA))
```
Checking the records

```{r}
tbl_transform_ped %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Animal")) %>% select(V12,V11)
tbl_transform_ped %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Parent")) %>% select(V12,V11)
```

Comparing the transformed pedigree records to the original records shows that birthdates are invalidated.

```{r}
tbl_ped_uni_id %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Animal")) %>% select(V12,V11)
tbl_ped_uni_id %>% inner_join(tbl_fail_mother_age, by = c("V12" = "Parent")) %>% select(V12,V11)
```

#### Function transform_check_parent_older_offspring

```{r}
(tbl_transform_ped <- transform_check_parent_older_offspring(ptbl_pedigree = tbl_ped_uni_id,
                                                            output_check = tbl_fail_mother_age,
                                                            pb_out = TRUE))
```

### Parents sex consistency
If the sex of the parent are inconsistent, the parent's sex field will be invalidate with NA.

```{r ParentSexInconsistency, eval=TRUE, include=FALSE, results='hide'}
sDataFileSex <- system.file(file.path("extdata","KLDAT_20170524_100_mother_sex.txt"), 
                             package = "PedigreeFromTvdData")
tbl_ped_sex <- laf_open_fwf_tvd_input(ps_input_file = sDataFileSex, pb_out = TRUE)
tbl_result_sex <- check_sex_tbl(ptblPedigree = tbl_ped_sex)
```

```{r}
(transform_check_sex_tbl(ptbl_pedigree = tbl_ped_sex,
                         output_check = tbl_result_sex,
                         pb_out = TRUE))
```




```{r}
```

